---
title: "MitH_Data_Preprocessing Module"
author: "Rohit"
output: html_document
  
---

```{r setup, include=FALSE}
rm(list=ls(all.names=T))
data_comp1<-read.table(sep ="," ,na.strings ="?" ,as.is = T,file='D:/MITH/Dataset/Test_data_model_grader.csv',header=T)
head(data_comp)
str(data_comp)
summary(data_comp[d])
library(tidyverse)
data_comp%>%filter(istrain==1)->data_filter
d<-c('RowID','istrain','datacollected','EmployeeCount','FirstJobDate','DateOfjoiningintheCurrentCompany','EmployeeID','StandardHours')
data_train<-data_filter[setdiff(colnames(data_filter),d)]
str(data_train)
data_comp%>%filter(istrain==0)->data_filter1
data_filter1%>%select(RowID)->RowID
data_test<-data_filter1[setdiff(colnames(data_filter1),d)]

write.table(x = data_train,file = "D:/MITH/Dataset/Train_data_model.csv",sep = ",")
write.table(x = data_test,file = "D:/MITH/Dataset/Test_data_model_grader.csv",sep = ",")
write.table(x = RowID,file = "D:/MITH/Dataset/Test_data_model_g_ID.csv",sep = ",")
```
## Reading the dataset and Basic Pre-processing
1) Read data using read.table
2) Checking initial structure and summary of the data set
3) Iterations with required changes in the read.table command
4) Separating the numerical variables and categorical variables using sapply function
5) Adjusting data between the numeric and categorical dataframes
6) Using apply function to convert to factors
7) merge function incase two data sets given
```{r}
rm(list=ls(all.names = T))
data_comp<-read.table(sep ="," ,na.strings ="?" ,as.is = T,file='D:/MITH/Dataset/Train_data_model.csv',header=T)
#data_comp1<-read.table()
#merge(x=,y=,by.x=,by.y= )
data_test<-read.table(sep ="," ,na.strings ="?" ,as.is = T,file='D:/MITH/Dataset/Test_data_model_grader.csv',header=T)

summary(data_comp[cat_num])
str(data_comp[cat_num])
```
##Required data conversions and transformations
1)Separating columns of factors & characters from numeric & integer
2)Discretizing certain variables
  1)YearsSinceLastPromotion
  2)No.ofWorkshopsAttended
  3)YearsInCurrentRole  

```{r}
##for Train_data
num_vars<-colnames(data_comp)[sapply(data_comp,FUN = class) %in% c('numeric','integer')]
cat_vars<-colnames(data_comp)[sapply(data_comp,FUN = class) %in% c('factor','character')]

#adjustment based on definitions

cat_num<-c('Joblevel','OfficeAmbienceRating','SelfMotivationRating','RelationshipSatisfaction','Education','WorkLifeBalance','Happynesslevelinjob','ESOps','PerformanceRating') 

install.packages('rlist',repos = "http://cran.us.r-project.org")
library(rlist)
cat_vars<-list.append(cat_vars,cat_num)

num_vars<-setdiff(num_vars,cat_num)

data_numeric<-data_comp[num_vars]
data_cat<-data_comp[cat_vars]

##For test_data
num_vars<-colnames(data_test)[sapply(data_test,FUN = class) %in% c('numeric','integer')]
cat_vars<-colnames(data_test)[sapply(data_test,FUN = class) %in% c('factor','character')]

#adjustment based on definitions

cat_num<-c('Joblevel','OfficeAmbienceRating','SelfMotivationRating','RelationshipSatisfaction','Education','WorkLifeBalance','Happynesslevelinjob','ESOps','PerformanceRating') 

install.packages('rlist',repos = "http://cran.us.r-project.org")
library(rlist)
cat_vars<-list.append(cat_vars,cat_num)

num_vars<-setdiff(num_vars,cat_num)

data_numeric_t<-data_test[num_vars]
data_cat_t<-data_test[cat_vars]
summary(data_comp1[disc])
##To be discretized
disc<-c('YearsSinceLastPromotion','No.ofWorkshopsAttended','YearsInCurrentRole')
disc_data<-data_numeric[,disc]
disc_data_t<-data_numeric_t[,disc]
library(infotheo)
##Checking if test and train data for selected variables have the same range
summary(disc_data)
summary(disc_data_t)

###Discretizing in both test and train datasets

x<-discretize(disc_data,disc = 'equalwidth',nbins = 3)
x_test<-discretize(disc_data_t,disc = 'equalwidth',nbins = 3)
data_numeric[,disc]<-x
data_numeric_t[,disc]<-x_test

##Reiterating numerical and categorical variables

cat_vars<-list.append(cat_vars,disc)
num_vars<-setdiff(num_vars,disc)

data_cat<-cbind.data.frame(data_cat,x)
data_cat_t<-cbind.data.frame(data_cat_t,x_test)
```
## Final conversion to factors
```{r}
str(data_cat)
##For training data set
data_cat[sapply(data_cat,class) %in% c('numeric','integer')]<-as.data.frame(apply(data_cat[sapply(data_cat,class) %in% c('numeric','integer')],MARGIN = 2,FUN = as.character))
data_cat<-as.data.frame(apply(data_cat,MARGIN = 2,FUN=as.factor))

##For testing dataset
str(data_cat_t)
data_cat_t[sapply(data_cat_t,class) %in% c('numeric','integer')]<-as.data.frame(apply(data_cat_t[sapply(data_cat_t,class) %in% c('numeric','integer')],MARGIN = 2,FUN = as.character))
data_cat_t<-as.data.frame(apply(data_cat_t,MARGIN = 2,FUN=as.factor))

## Removing the discretized values from numerics and adding to categorical

data_numeric[disc]=NULL
data_numeric_t[disc]=NULL


data_final_train<-as.data.frame(cbind(data_numeric,data_cat))
data_final_test<-as.data.frame(cbind(data_numeric_t,data_cat_t))
str(data_final_train)
data_final_test$ExtraTime=NULL
data_final_test$Over18=NULL
data_final_train$Over18=NULL
```
##Checking for NA values
```{r}
library(DMwR)
table(data_final_test$ExtraTime)
colSums(is.na(data_final_train))
colSums(is.na(data_final_test))


```

## Dummy variables and Imputation
```{r}
library(caret)
!colnames(data_final_train)%in%c('ExtraTime')

library(DMwR)
dummyvar<-dummyVars(~.,data_final_train[!colnames(data_final_train)%in%c('ExtraTime')])
str(data_final_train)
data_dumm_train<-as.data.frame(predict(dummyvar,data_final_train))


standardize<-preProcess(data_dumm_train[num_vars],method = c('center','scale'))
data_standard<-predict(standardize,data_dumm_train)
data_standard['target']<-data_final_train$ExtraTime
str(data_standard$target)
##For test_data

dummyvar<-dummyVars(~.,data_final_test)

data_dumm_test<-as.data.frame(predict(dummyvar,data_final_test))


standardize1<-preProcess(data_dumm_test[num_vars],method = c('center','scale'))
data_standard_t<-predict(standardize1,data_dumm_test)

summary(data_standard[num_vars])
summary(data_standard_t[num_vars])

```
## Write the model ready data

```{r}
write.table(x = data_standard,file = "D:/MITH/Dataset/cleaned/Train_data_cleaned.csv",sep = ",")
write.table(x = data_standard_t, "D:/MITH/Dataset/Test_data_cleaned.csv",sep = ",")

```